---
output: github_document
---


```{r}
knitr::opts_chunk$set(message=FALSE)
```

```{r setup}
library(neonstore)
library(neon4cast) # remotes::install_github("eco4cast/neon4cast", dep=TRUE)
library(tidyverse)
library(tsibble)
library(lubridate)
library(fable)
```

# A trivial forecast


```{r}
## uuid::UUIDgenerate()
iteration_id <- uuid::UUIDgenerate()  #"c2cb06c2-4bf3-4aa0-9690-1fdd098a7ac5"

forecast_date <- Sys.Date() - months(3)

```

## Access Target Data

```{r}
targets <-
  "https://data.ecoforecast.org/neon4cast-targets/beetles/beetles-targets.csv.gz" %>% 
  read_csv(show_col_types = FALSE) 

#targets %>% duplicates(index = datetime, key  = c(variable,site_id))

targets_ts <- targets %>% 
  as_tsibble(index = datetime, key = c(variable,site_id))
```


```{r}
past <-  targets_ts %>% filter(datetime < forecast_date) |>
  pivot_wider(names_from="variable", values_from="observation")
```

## Compute a forecast

```{r}
## Compute a simple mean/sd model per site... obviously silly given huge seasonal aspect
null_richness <- past  %>% 
  model(null = MEAN(richness)) %>%
  forecast(h = "1 year")

null_abundance <- past  %>%
  model(null = MEAN(abundance)) %>%
  forecast(h = "1 year")
```


## Adjust prediction to ISOweek


## Visualize the forecast

```{r}
first4 <- unique(null_richness$site_id)[1:4]

null_richness %>% filter(site_id %in% first4)  %>% 
  autoplot(past) + ggtitle("richness")
null_abundance %>% filter(site_id %in% first4)  %>% 
  autoplot(past) + ggtitle("abundance")
```

## EFI Formatting

EFI requires a flat-file format for forecasts that avoids the use of complex list columns.  
To convey uncertainty, forecasts must be expressed either by giving mean and standard deviation (for predictions that are normally distributed) or must express forecasts as an ensemble of replicate draws from forecast distribution.
The helper function `efi_format()` handles this transformation.

```{r}
## Combine richness and abundance forecasts
forecast_raw <- bind_rows(efi_format(null_richness), 
                            efi_format(null_abundance)) 
```
Dates must be expressed as the Monday of the week the sampling would be performed.  
This way don't need to predict the precise day that the trap is collected, all traps
collected in the same week will count as an observation made for that week.


```{r}
iso_week <- function(x){
  monday <- gsub("\\d$", "1", ISOweek::date2ISOweek(x) )
  ISOweek::ISOweek2date(monday)
  }
```


```{r}
forecast <- forecast_raw %>% mutate(datetime = iso_week(datetime) )
```

```{r}
scores <- neon4cast::score(forecast, targets)

scores |> filter(!is.na(observation))
```




```{r}
forecast_file <- glue::glue("{theme}-{date}-{team}.csv.gz",
                            theme = "beetles", 
                            date=forecast_date,
                            team = "cb_f1")
write_csv(forecast, forecast_file)
```


```{r}
neon4cast::forecast_output_validator(forecast_file)
```


```{r}
submit(forecast_file)
```
