---
output: github_document
---


```{r}
knitr::opts_chunk$set(message=FALSE)
```

```{r setup}
library(neonstore)
library(tidyverse)
library(tsibble)
library(fable)
library(feasts)
library(urca)
library(ISOweek)
library(neon4cast) # remotes::install_github("eco4cast/neon4cast", dep=TRUE)
library(lubridate)
#source("R/flat_format.R")
source("R/iso_week.R")
```

# A trivial ARIMA forecast


ARIMA models are commonly used to predict future values from historical values, ([FPP Ch 9](https://otexts.com/fpp3/arima.html)).  A simple ARIMA model does not use any external driver variables, though it is possible to include regression (e.g. see [FPP Ch 10 on Dynamic Models](https://otexts.com/fpp3/dynamic.html))


Our ARIMA model needs only one extra step, making implicit missing data into explicit missing values.  In both richness and abundance, we will treat gaps as missing data, not as zeros.

```{r}
team <- "cb_f2"
## uuid::UUIDgenerate()
iteration_id <- "b7a6f03d-0bac-43f3-9216-da0f7fe4f231"
```

## Access Target Data

```{r}
targets <-
  "https://data.ecoforecast.org/neon4cast-targets/beetles/beetles-targets.csv.gz" |> 
  read_csv(show_col_types = FALSE) 

#targets |> duplicates(index = datetime, key  = c(variable,site_id))

targets_ts <- targets |> 
  as_tsibble(index = datetime, key = c(variable,site_id))
```


```{r}
forecast_date <- Sys.Date() - months(3)
past <-  targets_ts |>
  filter(datetime < forecast_date)  |>
  pivot_wider(names_from="variable", values_from="observation")
```

## Compute a forecast

```{r}
gap_filled <- tsibble::fill_gaps(past)
arima_richness <- gap_filled  |> 
  model(arima = ARIMA(richness)) |>
  forecast(h = "1 year") |>
  efi_format()

arima_abundance <- gap_filled  |>
  model(arima = ARIMA(abundance)) |>
  forecast(h = "1 year") |>
  efi_format()

```

## EFI Formatting

```{r}
## Combine richness and abundance forecasts. drop the 'model' column
forecast <- bind_rows(arima_richness, arima_abundance)

```
Dates must be expressed as the Monday of the week the sampling would be performed.  
This way don't need to predict the precise day that the trap is collected, all traps
collected in the same week will count as an observation made for that week.


```{r}
forecast <- forecast |> mutate(datetime = iso_week(datetime) )
```


```{r}
forecast_file <- glue::glue("{theme}-{date}-{team}.csv.gz",
                            theme = "beetles", 
                            date=forecast_date,
                            team = team)
write_csv(forecast, forecast_file)
```


```{r}
neon4cast::forecast_output_validator(forecast_file)
```



```{r}
submit(forecast_file)
```
