---
output: github_document
---

This document contains draft notebook exploring potential different forecasts.

```{r}
knitr::opts_chunk$set(message=FALSE)
```

```{r setup}
library(neonstore)
library(neon4cast) # remotes::install_github("eco4cast/neon4cast")
library(tidyverse)
library(tsibble)
library(fable)
```

# A trivial forecast


```{r}
targets <-
  "https://data.ecoforecast.org/neon4cast-targets/beetles/beetles-targets.csv.gz" |> 
  read_csv(show_col_types = FALSE) 

#targets |> duplicates(index = datetime, key  = c(variable,site_id))

targets_ts <- targets |> 
  as_tsibble(index = datetime, key = c(variable,site_id))
```


```{r}
forecast_date <- Sys.Date() - months(12)


past <-  targets_ts |>
  filter(datetime < forecast_date)  |>
  pivot_wider(names_from="variable", values_from="observation")


future <- targets_ts |>
  filter(datetime >= forecast_date)  |>
  pivot_wider(names_from="variable", values_from="observation")
```

## Compute a forecast

```{r}
## Compute a simple mean/sd model per site... obviously silly given huge seasonal aspect
null_richness <- past  |> 
  model(null = MEAN(richness)) |>
  forecast(h = "1 year")

null_abundance <- past  |>
  model(null = MEAN(abundance)) |>
  forecast(h = "1 year")
```


## Visualize the forecast

```{r}
first4 <- unique(null_richness$site_id)[1:4]

null_richness |> filter(site_id %in% first4)  |> autoplot(past) + ggtitle("richness")
null_abundance |> filter(site_id %in% first4)  |> autoplot(targets) + ggtitle("abundance")
```

## Score the forecast

`fable` can compute CRPS, by default computes averages by site_id (`key`) and `.model` only.  
Optionally, we could include `by = c(".model", "site_id", "time")` to generate separate forecasts by time. 
(For some reason `fable`s internal method is much slower than the EFI code, though results are numerically equivalent). 
Note that fable can work across multiple models, (though we have only one in this case), but not across multiple target variables.
We must handle predictions of `richness` and `abundance` separately.
Recall that smaller CRPS scores are better.
  

```{r}
null_richness |>
  accuracy(future, list(crps = CRPS))
```

## EFI Formatting

EFI requires a flat-file format for forecasts that avoids the use of complex list columns.  
To convey uncertainty, forecasts must be expressed either by giving mean and standard deviation (for predictions that are normally distributed) or must express forecasts as an ensemble of replicate draws from forecast distribution.
The helper function `efi_format()` handles this transformation.

```{r}
## Combine richness and abundance forecasts.
null_forecast <- bind_rows(efi_format(null_richness), 
                            efi_format(null_abundance)) 
```

Score the forecast using EFI's internal method. By default, EFI's method reports the score every unique site-time combination (unique grouping variables).
It is easy to later average across times for a by-site score.

```{r}
scores_null <- neon4cast::score(null_forecast, targets) |> filter(!is.na(observation))
# average richness scores by site
summary_scores <- scores_null |> 
  mutate(month = lubridate::month(datetime,label=TRUE)) |>
  group_by(variable,month) |>
  summarise(crps = mean(crps, na.rm=TRUE),
            logs = mean(logs, na.rm=TRUE),
            .groups = "drop") |>
  pivot_longer(c(crps, logs), names_to="metric", values_to="score")

summary_scores |>
  ggplot(aes(month, score)) + geom_col() +
  facet_grid(variable ~ metric, scales = "free_y")
```





